k-Space Domain Parallel Transmit Pulse Design

### 2020-08-31 Manuscript/Theory.tex
wag theory fig 1 edits

\par The columns of $\bm{W}$ can be solved independently, 
based on the observation that if the desired k-space pattern is a delta function at a k-space location 
$\vec{k}_j$ ($1\leq j \leq N_s$), 
then only the $j$-th column $\bm{w}_j$ of $\bm{W}$ is required to relate 
the desired k-space pattern to the desired RF pulse. 
In this case, the vector $\bm{w}_j$ is identical to the desired RF. 
Therefore, the problem of finding the $j$-th column $\bm{w}_j$ of $\bm{W}$ can be restated as: 
what should the RF pulse samples be to generate a unit delta function at the location $\vec{k}_j$ in k-space, 
and zeros elsewhere? 
Mathematically, the relationship that should be satisfied by the weights in column $\bm{w}_j$  can be expressed as:   
\begin{equation}\label{eq:ScalerForm}
\delta(\vec{k}-\vec{k}_j)=\sum_{i=1}^{N_t}\sum_{c=1}^{N_c} w_c(\vec{k}_i) s_c(\vec{k}-\vec{k}_i)
\end{equation}
where $i$ indexes time points in the pulses' excitation k-space trajectory, 
and $c$ indexes channels. 
$w_c(\vec{k}_i)$ is the $((c-1)N_t+i)$-th element of the $\bm{w}_j$ vector,
and $s_c(\vec{k}-\vec{k}_i)$ is the Fourier transform of channel-$c$'s $B_1^+$ map, 
shifted to be centered at excitation k-space location $\vec{k}_i$. 
Equation \ref{eq:ScalerForm} can be restated in matrix-vector form as:                                       
\begin{equation}\label{eq:MatrixForm}
\bm{Sw}_j=\bm{\delta}_j,
\end{equation}
where $\bm{S}$ is a $N_s \times N_c  N_t$ 
matrix containing values of $s_c(\vec{k}-\vec{k}_i)$ for all $c$ and $i$,
and $\bm{\delta}_j$ is a vector containing a one in the entry corresponding to excitation k-space location
$\vec{k}_j$, and zeros in all other entries. 
The regularized pseudoinverse solution for $\bm{w}_j$ is:                      
\begin{equation}\label{eq:Solution}
\bm{w}_j=\left( \bm{S}^{H} \bm{S} + \lambda \bm{I} \right) ^{-1} \bm{s}_j^{H}
\end{equation}
where $\bm{s}_j^{H}$ is the conjugate transpose of the $j$-th row of $\bm{S}$.

\par If all points on the excitation k-space trajectory are considered in Equation \ref{eq:ScalerForm},
then there is no reduction in computation compared to a conventional pseudoinverse-based pulse design.
However, because $B_1^+$ maps are compact in excitation k-space and localized near DC,
only a small number ($\ll N_t$) of trajectory points near the target location $\vec{k}_j$ can contribute significantly
to the excitation pattern at $\vec{k}_j$ or its neighbors, 
and therefore need to be considered in the subproblem;
all other trajectory points can be ignored.
This means we can modify Equation \ref{eq:ScalerForm} to:
\begin{equation}
\delta(\vec{k}-\vec{k}_j) = \sum_{\vec{k}_i \in \mathbb{N}}\sum_{c=1}^{N_c} w_c(\vec{k}_i) s_c(\vec{k}-\vec{k}_i),
\end{equation} 
where $\mathbb{N}$ is the set of trajectory points in the vicinity of $\vec{k}_j$. 
In this work we define `inclusion width' as the distance, 
in cycles per excitation field-of-view (cycles/FOV), 
from each target location $\vec{k}_j$ within which which all trajectory points are accrued into $\mathbb{N}$.
This reduced set of trajectory points reduces the column dimension of $\bm{S}$
and both dimensions of $\bm{S}^H\bm{S}$,
and implies that the design matrix $\bm{W}$ is sparse.
Figure \ref{fig:Patch}a illustrates a target point $\vec{k}_j$ for a two-dimensional spiral pulse design, 
and the nearby points on the spiral trajectory that need to be included in the solution for $\bm{w}_j$
for an inclusion width of 4 cycles/FOV.

@@

\begin{figure}
	\centering
	\includegraphics[width=6cm]{kspace_PTX_Patch}
	\caption{\textcolor{red}{WAG: I think you should make two subfigures here: one that illustrates a single target location, and one that illustrates a patch of target locations. Also you should annotate the different widths, in the figures, with arrows. Also add a legend for the dots} 
	Illustration of one instance within the palatalization of a 2D k-space domain design problem. The totally design problem is divided into patches of width 4 cycle/FOV, and the 16 yellow points are the target points in this instance. With inclusion width of 4 cycle/FOV, the 46 green excitation trajectory points within the 12 cycle/FOV wide green square (4 cycle/FOV extended outside of the patch) are considered in this instance. The excitation within the 20 cycle/FOV wide red patch is constrained to zero except the point where the delta function is.}
	\label{fig:Patch}
\end{figure}

@@

\subsection*{Off-Resonance Compensation}
% In order to model the off-resonance into our k-space domain design, we adapted a fast implementation of the off-resonance compensation with time segmentation approximation and weighted least-squares interpolators \cite{fessler2005toeplitz}. 
Off-resonance compensation is common in spatial-domain parallel pulse designs \cite{Grissom:2006:MRM}, 
but has not been previously implemented in a k-space-domain design.
Spatial domain designs based on non-uniform fast Fourier transforms make use of an approximate time-segmented
model for the off-resonance term \cite{fessler2005toeplitz},
which decouples the space- and time-dependence of off-resonant phase accrual into the product of 
space- and time-dependent terms, which are applied on either side of the non-uniform fast Fourier transform operator. 
Here we show that a time-segmented off-resonance model can be incorporated in the proposed k-space-domain algorithm. 
Mathematically, the time-segmented model is given by:
\begin{equation}
e^{\imath \Delta\omega_j (t_i-T) }\approx\sum_{l=1}^{L} g_l(t_i) e^{\imath \Delta\omega_j \breve{t}_l},
\end{equation}
where $\Delta\omega_j$ is the off-resonance frequency at location $\vec{x}_j$, 
$i$ indexes time points in the pulse, 
$T$ is the pulse duration,
and $l$ indexes the $L$ time segments, where $L$ is typically between 4 and 8. 
$g_{il}=b_{l}(t_i)$ are the temporal interpolators that multiply into the RF samples,
and $h_{lj}=e^{\imath \Delta\omega_j\breve{t}_l}$ are the off-resonance phase shifts 
applied to the excitation patterns produced by the RF pulse time segments centered at time points $\{\breve{t}_l: l = 1, \dots, L\}$.
We can incorporate the time-segmented off-resonance model into Equation \ref{eq:ScalerForm} as:
\begin{equation}
	\delta(\vec{k}-\vec{k}_j)=\sum_{l = 1}^L \sum_{i=1}^{N_t}\sum_{c=1}^{N_c} w_c(\vec{k}_i) g_l(t_i) \tilde{s}_{cl} (\vec{k}-\vec{k}_i), 
\end{equation}
where $\tilde{s}_{cl}(\vec{k})$ is the Fourier transform of the product of channel $c$'s $B_1^+$ map and time segment $l$'s 
off-resonance phase exponential. 
We can then write the equivalent of Equation \ref{eq:MatrixForm} with the off-resonance model:
\begin{equation}
	\sum_{l=1}^L \bm{S}_l \textrm{diag}(g_{il}) \bm{w}_j = \bm{\delta}_j,
\end{equation} 
where the matrix $\bm{S}_l$ is now formed from the $N_c  L$ Fourier transforms of the $B_1^+$ map and off-resonance
phase exponential products.  
Efficient construction of the $\bm{S}^{H}\bm{S}$ matrix follows much the same procedure as without
off-resonance except that with off-resonance,
the elements of each sub-matrix $\bm{S}_i^{H}\bm{S}_j$ in Equation \ref{eq:SHS_blocks}
are interpolated from $L(L+1)/2$ Fourier transformed maps,
and the interpolated values are weighted by the temporal interpolators for each pair of multiplied time segments before they 
are summed.
Compared to constructing $\bm{S}^H\bm{S}$ without off-resonance, 
taking off-resonance into account increases the total number of FFT operations form $N_c(N_c+1)/2$ to $N_cL(N_cL+1)/2$, 
and increases the number of interpolations needed for each sub-matrix element from 1 to $L(L+1)/2$.  
However, we note that only adjacent time segments with overlapping temporal interpolators at a given time point
need to be considered when calculating the elements of $\bm{S}^H\bm{S}$ corresponding to that time point.

>>>>

\par The columns of $\bm{W}$ can be solved independently, 
based on the observation that if the desired k-space pattern is a delta function at a k-space location 
$\vec{k}_j$ ($1\leq j \leq N_s$), 
then only the $j$-th column $\bm{w}_j$ of $\bm{W}$ is required to relate 
the desired k-space pattern to the desired RF pulse. 
In this case, the vector $\bm{w}_j$ is identical to the desired RF. 
Therefore, the problem of finding the $j$-th column $\bm{w}_j$ of $\bm{W}$ can be restated as: 
what should the RF pulse samples be to generate a unit delta function at the location $\vec{k}_j$ in k-space, 
and zeros elsewhere? 
Mathematically, the relationship that should be satisfied by the weights in column $\bm{w}_j$  can be expressed as:   
\begin{equation}\label{eq:ScalerForm}
\delta(\vec{k}-\vec{k}_j)=\sum_{i=1}^{N_t}\sum_{c=1}^{N_c} w_c(\vec{k}_i) s_c(\vec{k}-\vec{k}_i),
\end{equation}
where $i$ indexes time points in the pulses' excitation k-space trajectory, 
and $c$ indexes channels. 
$w_c(\vec{k}_i)$ is the $((c-1)N_t+i)$-th element of the $\bm{w}_j$ vector,
and $s_c(\vec{k}-\vec{k}_i)$ is the Fourier transform of channel-$c$'s $B_1^+$ map, 
shifted to be centered at excitation k-space location $\vec{k}_i$. 
Equation \ref{eq:ScalerForm} can be restated in matrix-vector form as:                                       
\begin{equation}\label{eq:MatrixForm}
\bm{Sw}_j=\bm{\delta}_j,
\end{equation}
where $\bm{S}$ is a $N_s \times N_c  N_t$ 
matrix containing values of $s_c(\vec{k}-\vec{k}_i)$ for all $c$ and $i$,
and $\bm{\delta}_j$ is a vector containing a one in the entry corresponding to excitation k-space location
$\vec{k}_j$, and zeros in all other entries. 
The regularized pseudoinverse solution for $\bm{w}_j$ is:                      
\begin{equation}\label{eq:Solution}
\bm{w}_j=\left( \bm{S}^{H} \bm{S} + \lambda \bm{I} \right) ^{-1} \bm{s}_j^{H}
\end{equation}
where $\bm{s}_j^{H}$ is the conjugate transpose of the $j$-th row of $\bm{S}$.

\par If all points on the excitation k-space trajectory are considered in Equation \ref{eq:ScalerForm},
then there is no reduction in computation compared to a conventional pseudoinverse-based pulse design.
However, because $B_1^+$ maps are compact in excitation k-space and localized near DC,
only a small number ($\ll N_t$) of trajectory points near the target location $\vec{k}_j$ can contribute significantly
to the excitation pattern at $\vec{k}_j$ or its neighbors, 
and therefore need to be considered in the subproblem;
all other trajectory points can be ignored.
This means Equation \ref{eq:ScalerForm} can be modified to:
\begin{equation}
\delta(\vec{k}-\vec{k}_j) = \sum_{\vec{k}_i \in \mathbb{N}}\sum_{c=1}^{N_c} w_c(\vec{k}_i) s_c(\vec{k}-\vec{k}_i),
\end{equation} 
where $\mathbb{N}$ is the set of trajectory points in the vicinity of $\vec{k}_j$. 
In this work `inclusion width' is defined as the distance, 
in cycles per excitation field-of-view (cycles/FOV), 
from each target location $\vec{k}_j$ within which which all trajectory points are accrued into $\mathbb{N}$.
This reduced set of trajectory points reduces the column dimension of $\bm{S}$
and both dimensions of $\bm{S}^H\bm{S}$,
and implies that the design matrix $\bm{W}$ is sparse.
Figure \ref{fig:Patch}a illustrates a target point $\vec{k}_j$ for a two-dimensional pulse design, 
and the nearby points on the trajectory that need to be included in the solution for $\bm{w}_j$
for an inclusion width of 4 cycles/FOV.

@@

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{kspace_PTX_Patch}
	\caption{(a) A target point $\vec{k}_j$ and the nearby trajectory points that are included when calculating 
	RF pulse weight contributions from this point.
	Inclusion width is defined as the distance from each target location $\vec{k}_j$ within which 
	which all trajectory points are considered when solving for the $j$-th column of the design ($\bm{W}$) matrix.
	(b) Calculation of the $\bm{W}$ matrix can be performed column-by-column for each target location $\vec{k}_j$,
	or patches of target points can be solved for jointly. 
	Here the 16 columns of $\bm{W}$ for a 4 $\times$ 4 patch of target points are solved together,
	and an inclusion width of 4 cycles/FOV dictates that all trajectory points within a 12 $\times$ 12 region
	centered on the patch are considered in the calculation.}
	\label{fig:Patch}
\end{figure}

@@

\subsection*{Off-Resonance Compensation}
% In order to model the off-resonance into our k-space domain design, we adapted a fast implementation of the off-resonance compensation with time segmentation approximation and weighted least-squares interpolators \cite{fessler2005toeplitz}. 
Off-resonance compensation is common in spatial-domain parallel pulse designs \cite{Grissom:2006:MRM}, 
but has not been previously implemented in a k-space-domain design.
Spatial domain designs based on non-uniform fast Fourier transforms make use of an approximate time-segmented
model for the off-resonance term \cite{fessler2005toeplitz},
which decouples the space- and time-dependence of off-resonant phase accrual into the product of 
space- and time-dependent terms, which are applied on either side of the non-uniform fast Fourier transform operator. 
Here we show that a time-segmented off-resonance model can be incorporated in the proposed k-space-domain algorithm. 
Mathematically, the time-segmented model is given by:
\begin{equation}
e^{\imath \Delta\omega_j (t_i-T) }\approx\sum_{l=1}^{L} g_l(t_i) e^{\imath \Delta\omega_j \breve{t}_l},
\end{equation}
where $\Delta\omega_j$ is the off-resonance frequency at location $\vec{x}_j$, 
$i$ indexes time points in the pulse, 
$T$ is the pulse duration,
and $l$ indexes the $L$ time segments, where $L$ is typically between 4 and 8. 
$g_{il}=b_{l}(t_i)$ are the temporal interpolators that multiply into the RF samples,
and $h_{lj}=e^{\imath \Delta\omega_j\breve{t}_l}$ are the off-resonance phase shifts 
applied to the excitation patterns produced by the RF pulse time segments centered at time points $\{\breve{t}_l: l = 1, \dots, L\}$.
The time-segmented off-resonance model can be incorporated into Equation \ref{eq:ScalerForm} as:
\begin{equation}
	\delta(\vec{k}-\vec{k}_j)=\sum_{l = 1}^L \sum_{i=1}^{N_t}\sum_{c=1}^{N_c} w_c(\vec{k}_i) g_l(t_i) \tilde{s}_{cl} (\vec{k}-\vec{k}_i), 
\end{equation}
where $\tilde{s}_{cl}(\vec{k})$ is the Fourier transform of the product of channel $c$'s $B_1^+$ map and time segment $l$'s 
off-resonance phase exponential. 
We can then write the equivalent of Equation \ref{eq:MatrixForm} with the off-resonance model:
\begin{equation}
	\sum_{l=1}^L \bm{S}_l \textrm{diag}(g_{il}) \bm{w}_j = \bm{\delta}_j,
\end{equation} 
where the matrix $\bm{S}_l$ is now formed from the $N_c  L$ Fourier transforms of the $B_1^+$ map and off-resonance
phase exponential products.  
Efficient construction of the $\bm{S}^{H}\bm{S}$ matrix follows much the same procedure as without
off-resonance except that with off-resonance,
the elements of each sub-matrix $\bm{S}_i^{H}\bm{S}_j$ in Equation \ref{eq:SHS_blocks}
are interpolated from $L(L+1)/2$ Fourier transformed maps,
and the interpolated values are weighted by the temporal interpolators for each pair of multiplied time segments before they 
are summed.
Compared to constructing $\bm{S}^H\bm{S}$ without off-resonance, 
taking off-resonance into account increases the total number of FFT operations form $N_c(N_c+1)/2$ to $N_cL(N_cL+1)/2$, 
and increases the number of interpolations needed for each sub-matrix element from 1 to $L(L+1)/2$.  
However, we note that only adjacent time segments with overlapping temporal interpolators at a given time point
need to be considered when calculating the elements of $\bm{S}^H\bm{S}$ corresponding to that time point.
